Formalizing Stack Safety as a Security Property

### 2022-12-03 submission5/noexceptions.tex

\section{The Setting}

We state our properties in terms of a generic machine subject to a few constraints,
but our examples and tests are focused on a RISC-V-like machine enhanced with PIPE,
a tag-based reference monitor. Importantly, such a machine has no explicit ``call''
instruction, only {\tt jal} instructions which may be used for that purpose or potentially
other purposes. Similarly, there is no ``return'', only {\tt jalr}, and so on.
We therefore define our properties under the assumption that we can distinguish the
steps that represent these abstract operations from those that do not---for instance,
by labeling instructions as calls, returns, etc.
Such a labeling scheme appears in our testing environment (Section [TBD]).% and is already necessary
%to correctly apply tag-based enforcement in practice (cite something about this.)

The set of special operations that are recognized in this way are termed
{\it security-relevant operations}. Our properties are phrased as a {\it security semantics},
which extends the underlying machine with additional context about its
security principals and which registers and memory they require to be secure.
This context evolves dynamically through the execution of security-relevant operations.
This notion of security may be phrased as
predicates on future states, e.g. assertions of the form
``When control returns to me...'', or relations on traces of future execution
(hyper-properties).

Unlike approaches that give a safe-by-construction semantics
for various security-relevant operations and then prove secure compilation to the underlying
enforcement mechanism, our machine is explicitly not safe unless the enforcement
is correctly instantiated; the security semantics gives us the means to judge whether
an enforcement scheme implements stack safety or not.

We will first give a detailed example of a security semantics for a simple setting
in which our security-relevant operations are restricted to making calls and returns, and allocating
private memory within the current function activation. We then extend this model
to the full system that we test, which features:
\begin{itemize}
\item Function calls and returns, with caller- and callee-saved registers
\item Allocation of strictly private and strictly public regions on the stack
\item Arguments spilled onto the stack and/or passed by reference
\item Exceptions
\item Tail-call Elimination
\end{itemize}

Finally, we give a property definition for stack safety in which (some) stack-allocated
objects are governed by a provenance-based capability system, accessible only to those
functions which have recieved a valid pointer to them.

@@

Otherwise, we do not assume that the code adheres to any particular
calling convention or implements a source language construct.
In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, of course, particularly on the behavior of
call and return sequences.

In general, it is impossible to distinguish buggy source code from an attacker; in
our examples we will identify one function or another as an attacker, but we do not
require any static division between trusted and untrusted code, and aim to protect
even buggy code.

This is a strong threat model, but hardware and timing attacks are out of scope,
and our properties are termination insensitive as a result of the enforcement mechanism.

@@

Our concurrency model is fairly
simplistic, assuming a fixed number of threads each with its own dedicated processor.
We model memory safe stack objects, but not a heap. Regions outside of
stacks can be used however the compiler likes, including as a heap, but no protection is
built in and our properties assume that if a pointer to a stack object is stored there,
it is permanently compromised.

@@

The building blocks of the machine are {\em words} and {\em registers}.
Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
and are drawn from the set \(\WORDS\).
Registers in the set \(\REGS\) are ranged over by \(\reg\), with the stack pointer
given the special name \(\SP\).
They are divided into sets of caller-saved and callee-saved registers;
Figure \ref{fig:RISCVregs} gives an example division for a RISC-V machine as will
appear in our examples and testing, along with the register names that will appear in
example code, and their default security class (explained below).

Along with the program counter, \(\PCname\), these are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS\).

@@

A {\em machine state} \(\mach\) is a map from components to a set \(\mathcal{V}\) of
values. A value \(v\) may be convertible to an address, written \(|v|\) (\(|\cdot|\)
being a partial function.) We write \(\mach[\component]\) to denote the value at
\(\component\) of \(\mach\) and \(\mach[v]\) to denote the value at \(|v|\) if
defined.

@@

Security-relevant operations consist of calls, returns, and allocations.
Calls identify the address of the call's target,
and the argument registers. Allocations and deallocations are phrased in terms of a pair
of an offset from the stack pointer and a size. Returns don't require any additional information.
%
\begin{align*}
  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{return} & \\
  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
\end{align*}
%
Security principals are function activations: at any given time, pending activations
have security requirements that are recorded in a call stack, and the current
activation tracks context information that will inform its requirements when
it makes a call. Each activation has a {\it view}
of the system that maps each state element to a {\it security class}:
it may be private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\object\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)). The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to (Figure \ref{fig:RISCVregs}): \(\sealed\) for callee-saved,
\(\unsealed\) for caller-saved, and \(\public\) otherwise. Elements of \(\mathcal{P}\),
such as the program counter, are always \(\public\). Even if they might deserve their
own protection, that protection is orthogonal to stack safety.

@@

%\section{Calls, Returns, and Private Allocations}
\section{Examples}

In this section we consider a simple setting with calls, returns, and private allocations,
on a RISC-V machine with a PIPE tagging system. For some type \(\mathcal{T}\) of tags,
\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\), with \(|(w,t)| = w\).
Figure \ref{fig:main} gives the C code and corresponding compiled RISC-V code
for a function {\tt main}. (Code is compiled using the standard RISC-V
calling conventions~\cite{??}; the stack grows towards lower addresses. We assume that
{\tt main} begins at address 0 and its callee {\tt f} at address 100.)

{\tt main} takes a secret as its argument and initializes a variable to contain
potentially sensitive data. We assume that observable events are stores
to a special address, called {\tt out}. 
{\tt main} calls another function {\tt f},
and afterward may decide to output its secret based on the contents of
its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
it should not---instead it will output the results of {\tt f}.

@@

Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
assembly code in Figure \ref{fig:f}.\footnote{We arrange that all the snippets
have the same length just to simplify some later explanations.}
%
In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
pointer, accesses {\tt secret}, and directly outputs it. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
can instead return that value so that {\tt main} stores it to {\tt out},
as in Figure \ref{subfig:indirect}.
%
Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive}
with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
(Figure \ref{subfig:integrity})!
Attacks of this kind do not violate {\tt main}'s confidentiality, but its
{\it integrity}.
In Figure \ref{subfig:WBCF}, the attacker arranges to return to the
wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
violating the program's {\it well-bracketed control flow} (WBCF.) Figure \ref{subfig:WBCF2}
also shows an attack that violates WBCF, this time by returning to the correct
program counter but with the stack pointer offset.

@@

Consider the attack in Figure \ref{subfig:WBCF}: the attacker adds
16 to the return address and then returns, thus bypassing the {\tt if}-test in the code and outputting
{\tt secret}. Before the call, the program counter is 16 and the stack pointer is 980,
so we define a predicate \(\ret\)
on states that should hold just after the return, namely, on states \(\mach\)
such that \(\mach[\PCname] = 20\) and \(\mach[\SP] = 980\).

We can identify the point just after the return (if a return occurs)
as the first state in which the pending call stack is smaller than it was
just after the call.
WBCF requires that if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
For nested calls, where the pending stack is initially larger, the same principle
applies: \(\ret ~ \mach\) must hold the next time the pending stack is the same size or smaller.

@@

In typical compiled code, WBCF is maintained by the compiler's call and return sequences.
As long as a function deallocates the same-sized frame that it allocates, and properly
saves and restores the return address, it should have WBCF---unless an error in the middle
subverts those sequences by overwriting one of the relevant elements, or by jumping into
the middle of the call or return sequence. We can therefore enforce WBCF by (1) ensuring that
the call and return sequences execute in order and (2) permitting only call and return
sequences to write to \(\SP\), {\tt ra}, and the address in memory where {\tt ra} is stored.

Let \(\mathcal{T} ::= \mathit{gen} | \mathit{seq} ~ n ~ m | \mathit{priv}\), named for general
execution, sequence, and privileged, respectively. Call and return sequences of length \(m\) are
tagged \(\mathit{seq} ~ 1 ~ m\) through \(\mathit{seq} ~ m ~ l\), and the micro-policy enforces
that they execute in order. The stack pointer and {\tt ra}
are tagged \(\mathit{priv}\). Only instructions tagged \(\mathit{seq} \dots\) may write to memory
or registers tagged \(\mathit{priv}\). Store instructions tagged \(\mathit{seq} \dots\) tag their
targets \(\mathit{priv}\).

@@

Like WBCF, stack integrity defines at the call a condition that must hold upon
return. This time the condition applies to all of the memory that a function has
allocated. In Figure \ref{fig:exec1} we see the lifecycle of an allocated frame:
upon allocation, the view labeled is \(\object\), and when a call is made, it instead
becomes \(\sealed\). Intuitively, the integrity of {\tt main}
is preserved if, when control returns to it, it can rely on any \(\sealed\) elements
to be identical to when it made the call.
%
Again, we need to know when a caller has been returned to,
and we use the same mechanism of checking the depth of the call stack.
%
In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
addresses 980 through 999 and callee-saved registers such as
the stack pointer. Note that callee-saved registers often change
during the call---but if the caller accesses them after the call, it should find them
restored to their prior value.

@@

Depth Isolation tracks, via a tag on the program counter,
the depth of the current activation. Stack frames are initialized and cleared, by writing
to every address in them, so that during a call at depth \(d\), the current activation's
frame is tagged with \(d\). Loads and stores targeting addresses so tagged failstop
if the program counter does not match. In this setting, at depth \(d\), all memory tagged
with a depth less than \(d\) corresponds to \(\sealed\) memory, and is inaccessible.
When its depth is reached via a return, it will have the same value, so integrity is
enforced.

On the other hand, Lazy Tagging and Clearing skips the (problematically linear time)
initialization and clearing steps, and only updates tags when memory is written to.
It allows a callee to overwrite sealed memory, but that write tags the corrupted memory
with the callee's identity. If the corrupted memory's owner (or any other activation) tries
to read it, only then does the micro-policy failstop.

@@

To characterize this situation, we define a state element to be \emph{irrelevant}
if its value can be replaced with any other value without changing the observable behavior
of the machine. More precisely, for a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if they
agree on the value of every element except those in \(\components\).
In the tagged environment, we additionally need the values in \(\components\)
to agree on tags, but not necessarily their data.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{variants.png}
  \caption{Integrity Violation}
  \label{fig:variant}
\end{figure}

In our example setting, the observation trace consists of the sequence of values written to {\tt out}.
The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the call at step 5,
assuming that {\tt sec} is 5. Similar to WBCF, we will define
\(\intProp\) as a predicate on states that holds if
all relevant sealed addresses in \(\mach\) are the same as after step 5.
We once again require that \(\intProp\) hold on the state following the matching return.
That state is reached by step 9. Although {\tt sensitive} has changed, it might be
irrelevant. Consider a variant state in which {\tt sensitive} has any other value, arbitrarily
choosing 43 (and keeping its tag, \(t_2'\), unchanged.) As execution continues after the return
from the original state, it passes the {\tt if}-test on {\tt sensitive}, whereas the execution
from the variant does not, resulting in differing outputs. Therefore {\tt sensitive} is relevant,
so \(\intProp\) does not hold, and integrity has indeed been violated.

@@

In the second case, the return states disagree on the value of \(\SP + 12\). But in neither
case has that value changed since the original variants. So the difference is inherited from
the original variation, and does not depend on a secret value either. This matters because,
following the return, the caller is allowed to access its stack frame, so those addresses
may influence future execution legitimately.

@@

Finally, we define caller confidentiality as the 
combination of internal and return-time confidentiality.
While we have separated the definitions of integrity and confidentiality,
in terms of tag-based enforcement, both Depth Isolation and Lazy Tagging and Clearing
enforce them.

@@

Others combine this protection with some degree of memory protection,
chiefly focusing on integrity. Skorstengaard et al. \cite{SkorstengaardSTK} characterize stack
safety as the combination of WBCF and {\it local state encapsulation} (LSE), and prove that their
capability-based enforcement satisfies both. LSE corresponds to integrity, plus the
confidentiality of the caller's frame, but neither the unallocated stack nor data left
behind in it by previous callees.
They offer a fully abstract overlay semantics, which formalizes WBCF and LSE
together as a safe-by-construction abstract machine. That machine does not allow a callee to
read or write its caller's frame, but might allow it to read data that is left behind below
the caller's frame, which is weaker than our full caller confidentiality and rules out
callee confidentiality entirely. Georges et al. \cite{Georges22:TempsDesCerises} correct
both limitations.

@@

\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
written \(\mach \approx_\components \nach\), if, for
all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).

>>>>

\section{Machine Model and Security Semantics}

  We state our properties in terms of a generic machine, making minimal assumptions
  on the machine architecture and the nature of its built-in safety enforcement mechanisms
  (if any).  Our machine model is meant to be a faithful description of the hardware,
  with a semantics that covers both ordinary instruction behavior and the behavior of
  any tags, capabilities, etc. that the machine might support. 
  This semantics is intentionally \emph{not} ``safe-by-construction;'' it gives
  no guarantee that the enforcement hardware is being used correctly to achieve desired
  security properties.

  On top of this machine model, we layer a {\it security semantics}, which extends the
  underlying machine with additional context about the identity of current and pending
  functions (which act as security principals) and about the registers and memory they require
  to be secure. This added context is purely notional;
  it does not affect the real machine. The security context
  evolves dynamically through the execution of {\it security-relevant operations},
  which include events like calls, returns, and frame manipulations.
  Our security properties are phrased in terms of this context, often as predicates
  on future states, e.g. of the form ``when control returns to the current function...'',
  or relations on traces of future execution (hyper-properties).

  The security-relevant operations typically correspond to underling machine instructions,
  but the correspondence may not be obvious just be inspecting the machine code.
  For example, in the tagged RISC-V machine we use in our examples and tests,
  calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
  and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
  instructions might also be used for other things. We therefore assume that there
  is some external mechanism for distinguishing which instructions are security-relevant;
  the most obvious technique is to add optional labels to instructions, and that is 
  what we do in this paper, but the theory does not depend on this choice.

  Our security semantics has similarities to the overlay semantics that have been proposed
  to characterize stack safety in the past~\cite{??}, but, like the basic machine
  semantics, it is not safe-by-construction. Rather, it gives us a means to \emph{define}
  the security properties that a correct enforcement scheme should implement, using
  some combination of hardware and software.

In the next section, we give a detailed example of a security semantics for a simple setting
in which our security-relevant operations are restricted to making calls and returns, and allocating
private memory within the current function activation. In later sections, we extend this model
to the full system that we test, which features:
\begin{itemize}
\item Function calls and returns, with caller- and callee-saved registers
\item Allocation of strictly private and strictly public regions on the stack
\item Arguments spilled onto the stack and/or passed by reference
\item Exceptions
\item Tail-call Elimination
\end{itemize}

Finally, we give a property definition for stack safety in which (some) stack-allocated
objects are governed by a provenance-based capability system, accessible only to those
functions which have received a valid pointer to them.

@@

Otherwise, we do not assume that the code adheres to any particular
calling convention or implements a source language construct.
In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, particularly on the behavior of
call and return sequences.

In general, it is impossible to distinguish buggy source code from an attacker; in
our examples we will identify one function or another as an attacker, but we do not
require any static division between trusted and untrusted code, and we aim to protect
even buggy code.

This is a strong threat model, but hardware and timing attacks are out of scope,
and our properties are termination insensitive as a result of how failstops are
modelled (see ~\ref{??}).

@@

Our concurrency model is fairly
simplistic, assuming a fixed number of threads each with its own dedicated processor.
\apt{Is concurrency still alive?}
We model memory-safe stack objects, but not a heap. Regions outside of
stacks can be used however the compiler likes, including as a heap, but no protection is
built in and our properties assume that if a pointer to a stack object is stored there,
it is permanently compromised.

@@

The building blocks of the machine are {\em words} and {\em registers}.
Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
and are drawn from the set \(\WORDS\).
Registers in the set \(\REGS\) are ranged over by \(\reg\), with the stack pointer
given the special name \(\SP\); 
some registers may be classified as caller-saved or callee-saved.
%
Along with the program counter, \(\PCname\), these are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \PCname | \WORDS | \REGS\).

@@

A {\em machine state} \(\mach\) is a map from elements to a set \(\mathcal{V}\) of
\emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\) (\(|\cdot|\).
The details of value structure depend on the specific machine being modelled;  
intuitively, the payload represents the part of the value that is relevant to
the behavior of the basic machine, while the rest of the value may contain
information relevant to a hardware enforcement mechanism (such as a tag).
We write \(\mach[\component]\) to denote the value at
\(\component\) of \(\mach\) and \(\mach[v]\) to denote the value at \(|v|\) if
defined.

@@

Security-relevant operations consist of calls, returns, and allocations.
Calls identify the address of the call's target,
and the argument registers. Allocations and deallocations are phrased in terms of a pair
of an offset from the stack pointer and a size. Returns don't require any additional information.
%
\begin{align*}
  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{return} & \\
  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
\end{align*}
%
Security principals are function activations: at any given time, pending activations
have security requirements that are recorded in a call stack, and the current
activation tracks context information that will inform its requirements when
it makes a call. Each activation has a {\it view}
of the system that maps each state element to a {\it security class}:
it may be private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\object\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)). The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to: % (Figure \ref{fig:RISCVregs}):
\(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.
%Elements of \(\mathcal{P}\),
%such as the program counter, are always \(\public\). Even if they might deserve their
%own protection, that protection is orthogonal to stack safety.

@@

%\section{Calls, Returns, and Private Allocations}
\section{Properties by Example}

In this section we consider a simple setting with calls, returns, and private allocations,
on a RISC-V machine with a PIPE tagging system. Here 
\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\),
for some type \(\mathcal{T}\) of tags to be elaborated later, with \(|(w,t)| = w\).
Figure \ref{fig:RISCVregs} gives an example division of the RISC-V registers into
security classes. 
Figure \ref{fig:main} gives the C code and corresponding compiled RISC-V code
for a function {\tt main}. (Code is compiled using the standard RISC-V
calling conventions~\cite{??}; the stack grows towards lower addresses. We assume that
{\tt main} begins at address 0 and its callee {\tt f} at address 100.)

{\tt main} takes a secret as its argument and initializes a variable to contain
potentially sensitive data. We assume that the only observable events are stores
to a special address, called {\tt out}. 
{\tt main} calls another function {\tt f},
and afterward may decide to output its secret based on the contents of
its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
it should not---instead it will output the results of {\tt f}.

@@

Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
assembly code in Figure \ref{fig:f}.\footnote{We arrange that all the snippets
have the same length just to simplify some later explanations.}
%
In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
pointer, accesses {\tt secret}, and directly outputs it. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
can instead return that value so that {\tt main} stores it to {\tt out},
as in Figure \ref{subfig:indirect}.
%
Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive}
with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
(Figure \ref{subfig:integrity})!
Attacks of this kind do not violate {\tt main}'s confidentiality, but its
{\it integrity}.
In Figure \ref{subfig:WBCF}, the attacker arranges to return to the
wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
violating the program's {\it well-bracketed control flow} (WBCF.) Figure \ref{subfig:WBCF2}
shows another attack that violates WBCF, this time by returning to the correct
program counter but with the stack pointer offset.

@@

Consider the attack in Figure \ref{subfig:WBCF}: the attacker adds
16 to the return address and then returns, thus bypassing the {\tt if}-test in the code and outputting
{\tt secret}. Before the call, the program counter is 16 and the stack pointer is 980,
so we define a predicate on states that should hold just after the return:
\(\ret\ \mach \triangleq \mach[\PCname] = 20 \wedge \mach[\SP] = 980\).
%
We can identify the point just after the return (if a return occurs)
as the first state in which the pending call stack is smaller than it was
just after the call.
WBCF requires that if \(\mach\) is the state at that point, then \(\ret ~ \mach\) holds.
%For nested calls, where the pending stack is initially larger, the same principle
%applies: \(\ret ~ \mach\) must hold the next time the pending stack is the same size or smaller.

@@

In typical compiled code, WBCF is maintained by the compiler's call and return sequences.
As long as a function deallocates the same-sized frame that it allocates, and properly
saves and restores the return address, it should have WBCF---unless an error in the middle
subverts those sequences by overwriting one of the relevant elements, or by jumping into
the middle of the call or return sequence. We can therefore enforce WBCF by (1) ensuring that
these ``blessed'' call and return sequences execute in order and (2) permitting only blessed
sequences to write to \(\SP\), {\tt ra}, and the address in memory where {\tt ra} is stored.

To perform this enforcement using tags, we can take
\(\mathcal{T} ::= \mathit{gen} | \mathit{bseq} ~ n ~ m | \mathit{prot}\), named for general
execution, blessed sequence, and protected, respectively. Blessed sequences of length \(m\) are
tagged \(\mathit{bseq} ~ 1 ~ m\) through \(\mathit{bseq} ~ m ~ l\), and the micro-policy enforces
that they execute in order. The stack pointer and {\tt ra}
are tagged \(\mathit{prot}\). Only instructions tagged \(\mathit{bseq}\) may write to memory
or registers tagged \(\mathit{prot}\). Store instructions tagged \(\mathit{bseq}\) tag their
targets \(\mathit{prot}\).

@@

Like WBCF, stack integrity defines a condition at the call that must hold upon
return. This time the condition applies to all of the memory that a function has
allocated. In Figure \ref{fig:exec1} we see the lifecycle of an allocated frame:
upon allocation, the view labels it \(\object\), and when a call is made, it instead
becomes \(\sealed\). Intuitively, the integrity of {\tt main}
is preserved if, when control returns to it, it can rely on any \(\sealed\) elements
to be identical to when it made the call.
%
Again, we need to know when a caller has been returned to,
and we use the same mechanism of checking the depth of the call stack.
%
In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
addresses 980 through 999 and callee-saved registers such as
the stack pointer. Note that callee-saved registers often change
during the call---but if the caller accesses them after the call, it should find them
restored to their prior value.

@@

Depth Isolation tracks the depth of the current activation via a tag on the program counter.
Stack frames are initialized and cleared, by writing
to every address in them, so that during a call at depth \(d\), the current activation's
frame is tagged with \(d\). Loads and stores targeting addresses in frames failstop
if the program counter tag does not match the frame tag. In this setting, at depth \(d\),
all memory tagged with a depth less than \(d\) corresponds to \(\sealed\) memory,
and should be indeed be inaccessible.

On the other hand, Lazy Tagging and Clearing skips the (inefficient)
initialization and clearing steps, and only updates tags when memory is written to.
It allows a callee to overwrite sealed memory, but that write tags the corrupted memory
with the callee's identity. Only if the corrupted memory's owner (or any other activation) tries
to read it will the micro-policy failstop.

@@

To characterize this situation in a property, we define a state element to be \emph{irrelevant}
if its value can be replaced with any other value without changing the observable behavior
of the machine. More precisely, for a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if they
agree on the value of every element not in \(\components\).
In the tagged environment, we additionally require that the values of
\emph{all} elements in \(\mach\) and \(\nach\) agree on their tags. 
\apt{Right?  This does not generalize so nicely, since we have not defined the
  non-payload part of values, but it will do ok here, I think.}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{variants.png}
  \caption{Integrity Violation\apt{Tags are too small to read (also in next figures)}}
  \label{fig:variant}
\end{figure}

In our example setting, the observation trace consists of the sequence of values written to {\tt out}.
The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the call at step 5,
assuming that {\tt sec} is 5. Similar to WBCF, we will define
\(\intProp\) as a predicate on states that holds if
all relevant sealed addresses in \(\mach\) are the same as after step 5.
We once again require that \(\intProp\) hold on the state following the matching return, 
which is reached by step 9. Here {\tt sensitive} has obviously changed, but is it relevant?
Consider a variant state in which {\tt sensitive} has any other value, arbitrarily
choosing 43 (and keeping its tag, \(t_2'\), unchanged.) As execution continues after the return
from the original state, it passes the {\tt if}-test on {\tt sensitive}, whereas the execution
from the variant does not, resulting in differing outputs. Therefore {\tt sensitive} \emph{is} relevant,
so \(\intProp\) does not hold, and integrity has indeed been violated.

@@

In the second case, the return states disagree on the value of \(\SP + 12\). But in neither
case has that value changed since the original variants. So the difference is inherited from
the original variation, and does not depend on a secret value either. This matters because,
following the return, the caller is allowed to access its stack frame, so those addresses
may legitmately influence future execution.

@@

Finally, we define caller confidentiality as the 
combination of internal and return-time confidentiality.
While we have separated the definitions of integrity and confidentiality,
in terms of tag-based enforcement, Depth Isolation and Lazy Tagging and Clearing
both enforce both properties.

@@

Others combine this protection with some degree of memory protection,
chiefly focusing on integrity. Skorstengaard et al. \cite{SkorstengaardSTK} characterize stack
safety as the combination of WBCF and {\it local state encapsulation} (LSE), and prove that their
capability-based enforcement satisfies both. LSE corresponds to integrity, plus the
confidentiality of the caller's frame, but not of the unallocated stack or data left
behind in it by previous callees.
They offer a fully abstract overlay semantics, which formalizes WBCF and LSE
together as a safe-by-construction abstract machine. That machine does not allow a callee to
read or write its caller's frame, but might allow it to read data that is left behind below
the caller's frame, which is weaker than our full caller confidentiality and rules out
callee confidentiality entirely. Georges et al. \cite{Georges22:TempsDesCerises} correct
both limitations.

@@

\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
written \(\mach \approx_\components \nach\), if, for
all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\).
\apt{Needs revision to discuss payloads/tags.}
